---
title: 'Example 1: Using sourceCPP()'
author: "James Henderson, PhD"
date: "June 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<a href='./Intro_to_Rcpp.html'>Back</a> to main page.

#### Example 1: Identifying Points in Regions

##### The task 
In this example, we identify which of eight regions each point in a scatter plot belongs to.

```{r,echo=TRUE,include=FALSE}
library(ggplot2); library(dplyr)
```

```{r,echo=FALSE}
x = c(1,0,-1,0,1,1,1,1); y = c(0,1,0,-1,0,0,0,0)
data.frame(x=x,y=y) %>%
  ggplot(aes(x=x,y=y)) + theme_minimal() + 
  geom_hline(aes(yintercept=0)) + geom_vline(aes(xintercept=0)) +
  geom_path() + xlim(c(-3,3)) + ylim(c(-3,3)) + 
  geom_label(x=c(1,-1,-1,1,.25,-.25,-.25,.25),y=c(1,1,-1,-1,.25,.25,-.25,-.25),label=1:8,fill='gold')
```

##### R functions
<a href='./examples/getRegion.R'>getRegion.R</a>

##### C++ function
<a href='./examples/getRegion.cpp'>getRegion.cpp</a>

```{r,EightRegions}
library(Rcpp)
sourceCpp('examples/getRegion.cpp')
source('examples/getRegion.R')
```

```{r,EightRegionCheck}
n = 1e3; x = rnorm(n,0,.5); y = rnorm(n,.1,1)
Regions = data.frame(
  x = x, y=y,
  R_loop = getRegion_loop(x,y),
  R_vec  = getRegion_vec(x,y),
  R_all  = getRegion_all(x,y),
  CPP    = getRegionCPP(x,y)
)
## Test that there are no discrepencies
with(Regions,!{any(R_loop!=R_vec) | any(R_vec!=R_all) | any(R_vec!=CPP)})
Regions %>% mutate(Region=factor(CPP)) %>% ggplot(aes(x=x,y=y,color=Region)) +
  geom_point() + theme_minimal()
```

```{r,EightRegionTiming}
library(microbenchmark)
print(microbenchmark(getRegion_loop(x,y),getRegion_vec(x,y),getRegion_all(x,y),
                     getRegionCPP(x,y)),digits=2)
```

#### Exercises

1. Write an R function to identify coordinates in the regions 
between the lines $y=-\frac{1}{2}x$ and $y=2x$ as labeled below:

```{r,echo=FALSE}
x = c(1,-1,-1,1); y=c(1,1,-1,-1)
data.frame(x=x,y=y) %>%
  ggplot(aes(x=x,y=y)) + theme_minimal() + 
  xlim(c(-3,3)) + ylim(c(-3,3)) + geom_abline(slope=c(-.5,2)) + 
  geom_label(label=1:4,fill='gold')
```

2. Write a **C++** function for the same task.

3. Compare the computation times of your implementation using `microbenchmark`.
How is this example different than the previous one?


<a href='./Intro_to_Rcpp.html'>Back</a> to main page.